---
title: "Basic Exploration of Data"
author: "Emily B., Sagma P., Bridget E."
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(ggplot2)
```

### Load Data

Importing 2-year and 4-year college datasets. 

```{r code1}
two_year <- read_csv("Data_2-year_v5.csv")
four_year <- read_csv("Data_4-year_v5.csv")
View(two_year)
View(four_year)
```

### Filtering Datasets as Needed

Filtering datasets for building some initial visualizations in exploring the datasets. 

```{r code2}
#Fitler to include only schools in Washington in both the two year and four year datasets.
two_year_WA <- filter(two_year, fips_ipeds=="Washington")
four_year_WA <- filter(four_year, fips_ipeds=="Washington")

#Filter to include only schools in Washington in only the year 2017 in both the two year and four year datasets. 
two_year_WA_2017 <- filter(two_year_WA, year=="2017")
four_year_WA_2017 <- filter(four_year_WA, year=="2017")

#Pivot longer by market percent.
two_year_WA_2017_mkt <- pivot_longer(two_year_WA_2017, cols=c(mkt_white, mkt_hispa, mkt_black, mkt_asian, mkt_amind, mkt_pacis, mkt_twora), names_to="Market", values_to="MarketPercent")

four_year_WA_2017_mkt <- pivot_longer(four_year_WA_2017, cols=c(mkt_white, mkt_hispa, mkt_black, mkt_asian, mkt_amind, mkt_pacis, mkt_twora), names_to="Market", values_to="MarketPercent")

#Filter two year and four year school dataframes to incldue only information on schools in the year 2017.
two_year_2017 <- filter(two_year, year=="2017")
four_year_2017 <- filter(four_year, year=="2017")

#Filter to only look at SPU data, and more specifically SPU data in 2017. 
spu <- filter(four_year, inst_name=="Seattle Pacific University")
spu_2017 <- filter(spu, year=="2017")

#Pivot SPU data longer by market percent.
spu_mkt <-pivot_longer(spu, cols=c(mkt_white, mkt_hispa, mkt_black, mkt_asian, mkt_amind, mkt_pacis, mkt_twora), names_to="Market", values_to="MarketPercent")

#Pivot SPU data longer by college percent.
spu_colrep <-pivot_longer(spu, cols=c(col_white, col_hispa, col_black, col_asian, col_amind, col_pacis, col_twora), names_to="RaceRep", values_to="CollegePercent")
```

### Comparative Vertical Boxplots for Washington Schools 2017 (White and Black Representation)

Interesting trend of under-representation of black students in four-year Washington "private more-selective" colleges in 2017. 
Interesting trend of under-representation of white students in all almost all two-year Washington colleges.

```{r code3}
#Create plot to show Black Representation Gap in Washington two year colleges (2017). 
ggplot(two_year_WA_2017, aes(x= inst_name, y= dif_black, fill=twocat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="Black Representaion in Washington \nTwo-Year Colleges 2017",
         y = "College Representation of Black Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Instutional Category")
       
#Create plot to show White Representation Gap in Washington two year colleges (2017). 
ggplot(two_year_WA_2017, aes(x= inst_name, y= dif_white, fill=twocat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="White Representaion in Washington \nTwo-Year Colleges 2017",
         y = "College Representation of Black Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Institutional Category")

#Create plot to show Black Representation Gap in Washington four year colleges (2017). 
ggplot(four_year_WA_2017, aes(x= inst_name, y= dif_black, fill=fourcat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="Black Representaion in Washington \nFour-Year Colleges 2017",
         y = "College Representation of White Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Institutional Category")

#Create plot to show White Representation Gap in Washington four year colleges (2017). 
ggplot(four_year_WA_2017, aes(x= inst_name, y= dif_white, fill=fourcat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="White Representaion in Washington \nFour-Year Colleges 2017",
         y = "College Representation of White Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Institutional Category")
```

### Comparative Vertical Boxplots for Washington Market Percentages 2017 

Spokane has comparatively higher white market percentage.
Biggest variation in market percent seems to be for black students in four year colleges 2017.
More variation in market percents for two-year WA colleges than four-year.

```{r code4}
#Create plot to show College Market percents (by race/ethnicity) for Washington two year colleges (2017).
ggplot(two_year_WA_2017_mkt, aes(x= inst_name, y= MarketPercent, fill= Market))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="College Market in Washington \nTwo-Year Colleges 2017",
         y = "College Market Percentages 2017",
         x = "College Institution")
         
#Create plot to show College Market percents (by race/ethnicity) for Washington four year colleges (2017).
ggplot(four_year_WA_2017_mkt, aes(x= inst_name, y= MarketPercent, fill= Market))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="College Market in Washington \nFour-Year Colleges 2017",
         y = "College Market Percentages 2017",
         x = "College Institution")
```

### Exploring Institution Category Counts over the Years

```{r code6}
#Create barchart to show institution types (Selectivity & Private/Public) for all two year schools.
ggplot(two_year, aes(x= twocat, fill=twocat))+
  geom_bar()+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="College Two-year Insitution Category Counts",
         y = "Counts",
         x = "College Institution Category",
         fill= "Institution Category")
         
#Create barchart to show institution types (Selectivity & Private/Public) for all four year schools.
ggplot(four_year, aes(x= fourcat, fill=fourcat))+
  geom_bar()+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="College Four-year Insitution Category Counts",
         y = "Counts",
         x = "College Institution Category",
         fill= "Institution Category")

```

### Exploring SPU Data

```{r code8}
#Create historgram to show SPU college market percents (based on race/ethnicity) over the years 2009-2017.
ggplot(spu_mkt, aes(x= Market, y=MarketPercent, fill=Market))+
  geom_histogram(stat="identity")+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="SPU College Market Percents over the Years",
         y = "Market Percents",
         x = "Market Race",
         fill= "Market Race")
         
#Create historgram to show SPU college enrolled percents (based on race/ethnicity) over the years 2009-2017.
ggplot(spu_colrep, aes(x= RaceRep, y=CollegePercent, fill=RaceRep))+
  geom_histogram(stat="identity")+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="SPU College Enrolled Percents over the Years",
         y = "College Percents",
         x = "College Race",
         fill= "College Race")
```
