---
title: "Basic Exploration of Data"
author: "Emily Bogle"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(ggplot2)
```

### Load Data

Importing 2-year and 4-year college datasets. 

```{r code1}
two_year <- read_csv("Data_2-year_v5.csv")
four_year <- read_csv("Data_4-year_v5.csv")
View(two_year)
View(four_year)
```

### Filtering Datasets as Needed

Filtering datasets for building some initial visualizations in exploring the datasets. 

```{r code2}
two_year_WA <- filter(two_year, fips_ipeds=="Washington")
four_year_WA <- filter(four_year, fips_ipeds=="Washington")

two_year_WA_2017 <- filter(two_year_WA, year=="2017")
four_year_WA_2017 <- filter(four_year_WA, year=="2017")

two_year_WA_2017_mkt <- pivot_longer(two_year_WA_2017, cols=c(mkt_white, mkt_hispa, mkt_black, mkt_asian, mkt_amind, mkt_pacis, mkt_twora), names_to="Market", values_to="MarketPercent")

four_year_WA_2017_mkt <- pivot_longer(four_year_WA_2017, cols=c(mkt_white, mkt_hispa, mkt_black, mkt_asian, mkt_amind, mkt_pacis, mkt_twora), names_to="Market", values_to="MarketPercent")

two_year_2017 <- filter(two_year, year=="2017")
four_year_2017 <- filter(four_year, year=="2017")

spu <- filter(four_year, inst_name=="Seattle Pacific University")
spu_2017 <- filter(spu, year=="2017")

spu_mkt <-pivot_longer(spu, cols=c(mkt_white, mkt_hispa, mkt_black, mkt_asian, mkt_amind, mkt_pacis, mkt_twora), names_to="Market", values_to="MarketPercent")

spu_colrep <-pivot_longer(spu, cols=c(col_white, col_hispa, col_black, col_asian, col_amind, col_pacis, col_twora), names_to="RaceRep", values_to="CollegePercent")
```

### Comparative Vertical Boxplots for Washington Schools 2017 (White and Black Representation)

Interesting trend of under-representation of black students in four-year Washington "private more-selective" colleges in 2017. 
Interesting trend of under-representation of white students in all almost all two-year Washington colleges.

```{r code3}
ggplot(two_year_WA_2017, aes(x= inst_name, y= dif_black, fill=twocat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="Black Representaion in Washington \nTwo-Year Colleges 2017",
         y = "College Representation of Black Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Instutional Category")

ggplot(two_year_WA_2017, aes(x= inst_name, y= dif_white, fill=twocat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="White Representaion in Washington \nTwo-Year Colleges 2017",
         y = "College Representation of Black Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Institutional Category")

ggplot(four_year_WA_2017, aes(x= inst_name, y= dif_black, fill=fourcat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="Black Representaion in Washington \nFour-Year Colleges 2017",
         y = "College Representation of White Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Institutional Category")

ggplot(four_year_WA_2017, aes(x= inst_name, y= dif_white, fill=fourcat))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="White Representaion in Washington \nFour-Year Colleges 2017",
         y = "College Representation of White Population 2017 \n(College Percent minus Market Percent)",
         x = "College Institution",
       fill="Institutional Category")
```

### Comparative Vertical Boxplots for Washington Market Percentages 2017 

Spokane has comparatively higher white market percentage.
Biggest variation in market percent seems to be for black students in four year colleges 2017.
More variation in market percents for two-year WA colleges than four-year.

```{r code4}
ggplot(two_year_WA_2017_mkt, aes(x= inst_name, y= MarketPercent, fill= Market))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="College Market in Washington \nTwo-Year Colleges 2017",
         y = "College Market Percentages 2017",
         x = "College Institution")

ggplot(four_year_WA_2017_mkt, aes(x= inst_name, y= MarketPercent, fill= Market))+
  geom_histogram(stat = "identity")+
  coord_flip()+
  labs(title="College Market in Washington \nFour-Year Colleges 2017",
         y = "College Market Percentages 2017",
         x = "College Institution")
```

### Exploring Institution Category Counts over the Years

```{r code6}
ggplot(two_year, aes(x= twocat, fill=twocat))+
  geom_bar()+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="College Two-year Insitution Category Counts",
         y = "Counts",
         x = "College Institution Category",
         fill= "Institution Category")

ggplot(four_year, aes(x= fourcat, fill=fourcat))+
  geom_bar()+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="College Four-year Insitution Category Counts",
         y = "Counts",
         x = "College Institution Category",
         fill= "Institution Category")

```

### Exploring SPU Data

```{r code8}
ggplot(spu_mkt, aes(x= Market, y=MarketPercent, fill=Market))+
  geom_histogram(stat="identity")+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="SPU College Market Percents over the Years",
         y = "Market Percents",
         x = "Market Race",
         fill= "Market Race")

ggplot(spu_colrep, aes(x= RaceRep, y=CollegePercent, fill=RaceRep))+
  geom_histogram(stat="identity")+
  facet_grid(~year)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="SPU College Enrolled Percents over the Years",
         y = "College Percents",
         x = "College Race",
         fill= "College Race")
```