---
title: "Lab Name"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(car)
library(HH)
library(leaps)
library(ggplot2)

Additiona_data <- read.csv("Additiona_data.csv")

data("Additiona_data")

```

```{r}
cor(Additiona_data)
```

### "number_applied" 
```{r}
model_1 <- lm((mag_diff) ~ number_applied, data=Additiona_data)
summary(m1)
ggplot(Additiona_data, aes(x=(number_applied), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" number_applied vs Manitute difference",
         y = " Magnitute difference",
         x = " number_applied)")


m1 <- lm(log(mag_diff) ~ number_applied, data=Additiona_data)
summary(m1)
ggplot(Additiona_data, aes(x= log10 (number_applied), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" number_applied vs Manitute difference",
         y = " Magnitute difference",
         x = " number_applied)")

plot(m1$resid ~ m1$fitted)
abline(h=0)
qqPlot(m1$resid)
```

###  "number_admitted"  
```{r}
m2 <- lm(log(mag_diff) ~ number_applied, data=Additiona_data)
summary(m2)
ggplot(Additiona_data, aes(x= log10 (number_admitted), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" number_admitted vs Manitute difference",
         y = " Magnitute difference",
         x = " number_admitted")

plot(m2$resid ~ m2$fitted)
abline(h=0)
qqPlot(m2$resid)
```

### "number_enrolled" 
```{r}
m3 <- lm(log(mag_diff) ~ number_enrolled, data=Additiona_data)
summary(m3)
ggplot(Additiona_data, aes(x= log10 (number_enrolled), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" number_enrolled vs Manitute difference",
         y = " Magnitute difference",
         x = " number_enrolled")

plot(m3$resid ~ m3$fitted)
abline(h=0)
qqPlot(m3$resid)
```

### "rev_total_current"
```{r}
m4 <- lm(log(mag_diff) ~ rev_total_current, data=Additiona_data)
summary(m4)
ggplot(Additiona_data, aes(x= log10 (rev_total_current), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" rev_total_current vs Manitute difference",
         y = " Magnitute difference",
         x = " rev_total_current")

plot(m4$resid ~ m4$fitted)
abline(h=0)
qqPlot(m4$resid)
```

### "sch_total_student_aid"
```{r}
m5 <- lm(log(mag_diff) ~ sch_total_student_aid, data=Additiona_data)
summary(m5)
ggplot(Additiona_data, aes(x= log10 (sch_total_student_aid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" sch_total_student_aid vs Manitute difference",
         y = " Magnitute difference",
         x = " sch_total_student_aid")

plot(m5$resid ~ m5$fitted)
abline(h=0)
qqPlot(m5$resid)
```

### "exp_instruc_total" 
```{r}
m6 <- lm(log(mag_diff) ~ exp_instruc_total, data=Additiona_data)
summary(m6)
ggplot(Additiona_data, aes(x= log10 (exp_instruc_total), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" exp_instruc_total vs Manitute difference",
         y = " Magnitute difference",
         x = " exp_instruc_total")

plot(m6$resid ~ m6$fitted)
abline(h=0)
qqPlot(m6$resid)
```

### "tuition_fees_published"  
```{r}
m7 <- lm(log(mag_diff) ~tuition_fees_published, data=Additiona_data)
summary(m6)
ggplot(Additiona_data, aes(x= log10 (tuition_fees_published), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" tuition_fees_published vs Manitute difference",
         y = " Magnitute difference",
         x = " tuition_fees_published")

plot(m7$resid ~ m7$fitted)
abline(h=0)
qqPlot(m7$resid)
```

### "tuition_fees_ft"  
```{r}
m8 <- lm(log(mag_diff) ~ tuition_fees_ft, data=Additiona_data)
summary(m6)
ggplot(Additiona_data, aes(x= log10 ( tuition_fees_ft), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" tuition_fees_ft vs Manitute difference",
         y = " Magnitute difference",
         x = "tuition_fees_ft")

plot(m8$resid ~ m8$fitted)
abline(h=0)
qqPlot(m8$resid)
```

###  "student_faculty_ratio" 
```{r}
m9 <- lm(log(mag_diff) ~ student_faculty_ratio, data=Additiona_data)
summary(m9)
ggplot(Additiona_data, aes(x= log10 (student_faculty_ratio), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" student_faculty_ratio vs Manitute difference",
         y = " Magnitute difference",
         x = "student_faculty_ratio")

plot(m9$resid ~ m9$fitted)
abline(h=0)
qqPlot(m9$resid)
```

### "room_board"  
```{r}
m10 <- lm(log(mag_diff) ~ room_board, data=Additiona_data)
summary(m10)
ggplot(Additiona_data, aes(x= log10 (room_board), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" room_board vs Manitute difference",
         y = " Magnitute difference",
         x = "room_board")

plot(m10$resid ~ m10$fitted)
abline(h=0)
qqPlot(m10$resid)
```

### "endow_total" 
```{r}
m11 <- lm(log(mag_diff) ~ endow_total, data=Additiona_data)
summary(m10)
ggplot(Additiona_data, aes(x= log10 (endow_total), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" endow_total vs Manitute difference",
         y = " Magnitute difference",
         x = "endow_total")

plot(m11$resid ~ m11$fitted)
abline(h=0)
qqPlot(m11$resid)
```

### "faminc_mean"   
```{r}
m12 <- lm(log(mag_diff) ~ faminc_mean, data=Additiona_data)
summary(m12)
ggplot(Additiona_data, aes(x= log10 (faminc_mean), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" faminc_mean vs Manitute difference",
         y = " Magnitute difference",
         x = "faminc_mean")

plot(m12$resid ~ m12$fitted)
abline(h=0)
qqPlot(m12$resid)
```

### "grant_recipients_unitid"   
```{r}
m13 <- lm(log(mag_diff) ~ grant_recipients_unitid, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (grant_recipients_unitid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" grant_recipients_unitid vs Manitute difference",
         y = " Magnitute difference",
         x = "grant_recipients_unitid")

plot(m13$resid ~ m13$fitted)
abline(h=0)
qqPlot(m13$resid)
```
### "faminc_mean"   
```{r}
m14 <- lm(log(mag_diff) ~ faminc_mean, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (faminc_mean), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" faminc_mean vs Manitute difference",
         y = " Magnitute difference",
         x = "faminc_mean")

plot(m14$resid ~ m14$fitted)
abline(h=0)
qqPlot(m14$resid)
```
### "value_grants_disbursed_unitid" 
```{r}
m15 <- lm(log(mag_diff) ~value_grants_disbursed_unitid, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (value_grants_disbursed_unitid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" value_grants_disbursed_unitid vs Manitute difference",
         y = " Magnitute difference",
         x = "value_grants_disbursed_unitid")

plot(m15$resid ~ m15$fitted)
abline(h=0)
qqPlot(m15$resid)
```

### "loan_recipients_unitid" 
```{r}
m16 <- lm(log(mag_diff) ~ loan_recipients_unitid, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (loan_recipients_unitid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title=" loan_recipients_unitid vs Manitute difference",
         y = " Magnitute difference",
         x = "loan_recipients_unitid")

plot(m16$resid ~ m16$fitted)
abline(h=0)
qqPlot(m16$resid)
```

### "value_loan_disbursements_unitid" 
```{r}
m17 <- lm(log(mag_diff) ~ value_loan_disbursements_unitid, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (value_loan_disbursements_unitid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title="value_loan_disbursements_unitid vs Manitute difference",
         y = " Magnitute difference",
         x = "value_loan_disbursements_unitid")

plot(m17$resid ~ m17$fitted)
abline(h=0)
qqPlot(m17$resid)
```

### "campus_award_recipients_unitid" 
```{r}
m18 <- lm(log(mag_diff) ~ campus_award_recipients_unitid, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (campus_award_recipients_unitid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title="campus_award_recipients_unitid vs Manitute difference",
         y = " Magnitute difference",
         x = "campus_award_recipients_unitid")

plot(m18$resid ~ m18$fitted)
abline(h=0)
qqPlot(m18$resid)
```

### "value_campus_disbursed_unitid"  
```{r}
m19 <- lm(log(mag_diff) ~ value_campus_disbursed_unitid, data=Additiona_data)
summary(m13)
ggplot(Additiona_data, aes(x= log10 (value_campus_disbursed_unitid), y= mag_diff ))+
  geom_point() +  geom_smooth() +
  labs(title="value_campus_disbursed_unitid vs Manitute difference",
         y = " Magnitute difference",
         x = "value_campus_disbursed_unitid")

plot(m19$resid ~ m19$fitted)
abline(h=0)
qqPlot(m19$resid)
```
