---
title: "Line of Best Fit Plotting"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(ggplot2)
library(dplyr)
```

### Import, Filter, and Mutate Data

Insert any text here.

```{r code1}
two_year <- read_csv("CollegeData_2-year.csv")
four_year <- read_csv("CollegeData_4-year.csv")

four_year_clean <- four_year %>%
                  select(year, unitid, inst_name, dif_white, dif_amind, dif_asian, dif_black,
                        dif_pacis, dif_twora, dif_hispa)

two_year_clean <- two_year %>%
                  select(year, unitid, inst_name, dif_white, dif_amind, dif_asian, dif_black,
                        dif_pacis, dif_twora, dif_hispa)

four_year_md_smsd <- four_year_clean %>%
                  rowwise () %>%
                  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                                            abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                                            abs(dif_twora), na.rm = T)) %>%
                  mutate(se = sum((dif_white)^2, (dif_hispa)^2, (dif_black)^2, 
                              (dif_asian)^2, (dif_amind)^2, (dif_pacis)^2,
                              (dif_twora)^2, na.rm = T)) %>%
                  mutate(smsd = sqrt(1/7*se)) %>%
                  mutate(missing_racial_data = is.na(dif_white) + is.na(abs(dif_hispa)) 
                                              + is.na(abs(dif_black)) + is.na(abs(dif_asian)) 
                                              + is.na(abs(dif_amind)) + is.na(abs(dif_pacis))
                                              + is.na(abs(dif_twora))) %>%
                  mutate(missing_racial_data = as.character(missing_racial_data))

two_year_md_smsd <- two_year_clean %>%
                  rowwise () %>%
                  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                                            abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                                            abs(dif_twora), na.rm = T)) %>%
                  mutate(se = sum((dif_white)^2, (dif_hispa)^2, (dif_black)^2, 
                              (dif_asian)^2, (dif_amind)^2, (dif_pacis)^2,
                              (dif_twora)^2, na.rm = T)) %>%
                  mutate(smsd = sqrt(1/7*se)) %>%
                  mutate(missing_racial_data = is.na(dif_white) + is.na(abs(dif_hispa)) 
                                              + is.na(abs(dif_black)) + is.na(abs(dif_asian)) 
                                              + is.na(abs(dif_amind)) + is.na(abs(dif_pacis))
                                              + is.na(abs(dif_twora))) %>%
                  mutate(missing_racial_data = as.character(missing_racial_data))
```


### Plotting Line of Best Fit Graphs (Diff Magnitudes)

This code can be used with any institution. I have sample code below for SPU. Make necessary substitutions to look at other individual schools. 

```{r code2}
#Subset data to given school of interest
plot_data <- four_year_md_smsd %>%
                      filter(unitid == "455770")

##Make individual plots for any given school's data
ggplot(data = plot_data) +
  geom_point(mapping=aes(x=year, y=mag_diff, color=missing_racial_data)) +
  geom_smooth(mapping=aes(x=year, y=mag_diff), method = 'lm') +
  labs(title = "Total Magnitudes of Racial Representative Differnces \n Providence Christian College",
       x = "Years", y= "Total Magnitued of Racial Difference Percentages",
       color = "Missing Racial Data") +
       scale_x_continuous(breaks=seq(2009,2017,by=1))
```

### Comparison Plots Across Multiple Schools (Magnitude Difference Metric)


```{r code8}
plot_data <- four_year_md_smsd %>%
                      filter(unitid == "227429"| unitid == "198419"| unitid == "101480")

##Make individual plots for any given school's data
ggplot(data = plot_data, aes(x=year, y=mag_diff, color= inst_name))+
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Total Magnitudes of Racial Representative Differnces Across Schools",
       x = "Years", y= "Total Magnitued of Racial Difference Percentages",
       color="4-Year School (Rank)") +
       scale_x_continuous(breaks=seq(2009,2017,by=1))
```

### Plotting Line of Best Fit Graphs (sqrt-Mean Squared Differences)

This code can be used with any institution. I have sample code below for SPU. Make necessary substitutions to look at other individual schools. 

```{r code2}
#Subset data to given school of interest
SPU_smsd <- four_year_md_smsd %>%
                      filter(unitid == )

##Make individual plots for any given school's data
ggplot(data = SPU_smsd) +
  geom_point(mapping=aes(x=year, y=smsd, color=missing_racial_data)) +
  geom_smooth(mapping=aes(x=year, y=smsd), method = 'lm') +
  labs(title = "Total Magnitudes of Racial Representative Differnces",
       x = "Years", y= "Total Magnitued of Racial Difference Percentages",
       color="Missing Racial Data") +
       scale_x_continuous(breaks=seq(2009,2017,by=1))
```
