---
title: "Exploring Relationships"
author: ""
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(corrplot)
```

### Importing and Wrangling Data for Race Corrplot

```{r code1}
#Load four year colleges cvs
four_year <- read_csv("CollegeData_4-year.csv")

#Replace any missing N/A racial data in any given year for any given school with average of other years data
four_year <- four_year %>%
  mutate_at(vars(col_white, mkt_white, dif_white,
                 col_hispa, mkt_hispa, dif_hispa,
                 col_black, mkt_black, dif_black,
                 col_asian, mkt_asian, dif_asian,
                 col_amind, mkt_amind, dif_amind,
                 col_pacis, mkt_pacis, dif_pacis,
                 col_twora, mkt_twora, dif_twora), 
            ~replace_na(.,mean(., na.rm = TRUE)))


four_year_clean <- four_year %>%
                  select(unitid, year, fips_ipeds, inst_name, dif_white, dif_amind, dif_asian, dif_black,
                        dif_pacis, dif_twora, dif_hispa)

#NOTE: this will take a second to complete, green bar is good!, wait until it disappears 
four_year_md <- four_year_clean %>%
                  rowwise () %>%
                  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                                            abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                                            abs(dif_twora), na.rm = T))

###Cutting out schools with two or more years of data missing
four_year_keeper_list<- four_year_md %>%
                    group_by(unitid) %>%
                    summarize(count=n()) %>%
                    filter(count == 9 | count == 8)

four_year_md_cut <- four_year_md %>%
                    filter(unitid %in% four_year_keeper_list$unitid)

four_year_join <- four_year_md_cut %>%
          group_by(unitid) %>%
          summarise(avgdif_white = mean(dif_white), 
                    avgdif_amind = mean(dif_amind),
                    avgdif_asian = mean(dif_asian),
                    avgdif_black = mean(dif_black),
                    avgdif_pacis = mean(dif_pacis),
                    avgdif_twora = mean(dif_twora),
                    avgdif_hispa = mean(dif_hispa),
                    avgdif_magdif = (mean(mag_diff)))

four_year_final_md <- four_year_md_cut %>%   
                  group_by(unitid) %>%
                  do(linmod = lm( mag_diff ~ year, data = .)) %>%
                  rowwise () %>%
                  mutate(slope = linmod$coefficients[2]) %>%
                  arrange(slope)

four_year_md_join <- left_join(four_year_join, four_year_final_md, by=c('unitid'='unitid')) %>%
                  select(unitid, linmod, slope,
                    avgdif_white, 
                    avgdif_amind,
                    avgdif_asian,
                    avgdif_black,
                    avgdif_pacis,
                    avgdif_twora,
                    avgdif_hispa,
                    avgdif_magdif)
four_year_md_final_race <- four_year_md_join %>%
                    arrange(slope)

#For four year schools, create rank_md variable to store rank of md slopes across schools.
four_year_md_final_race$rank_md<- rank(four_year_final_md$slope)

four_year_race_corr <- four_year_md_final_race %>%
                      select(avgdif_white, 
                    avgdif_amind,
                    avgdif_asian,
                    avgdif_black,
                    avgdif_pacis,
                    avgdif_twora,
                    avgdif_hispa,
                    avgdif_magdif,
                    rank_md)
```

### Correlation Matrix for Race and Rank

```{r code2}
M1 = cor(four_year_race_corr)
corrplot(M1, method ='number')
```

### Wrangling Data for Institution Type Corrpot


```{r code3}
four_year_inst_type_join <- four_year %>%
  select(unitid, public, private, forprofit, selective, more_selective, non_selective) %>%
  group_by(unitid) %>%
  summarise(public = mean(public),
            private = mean(private),
            forprofit = mean(forprofit),
            selective = mean(selective),
            more_selective = mean(more_selective),
            non_selective = mean(non_selective))

four_year_inst_type_corr <- left_join(four_year_md_final, four_year_inst_type_join, by=c('unitid'='unitid')) %>%
  select(public, private, forprofit, selective, more_selective, non_selective, rank_md)

```

### Correlation Matrix for Inst. Type and Rank

```{r code789}
M2 = cor(four_year_inst_type_corr)
corrplot(M2, method ='number')
```

### Wrangling Data for Institution Type and Mag_Diff Corrpot


```{r code3}
four_year_type_magdiff <- four_year %>%
  select(unitid, public, private, forprofit, selective, non_selective, more_selective, dif_white, 
         dif_amind, dif_asian, dif_black, dif_pacis, dif_twora, dif_hispa) %>%
  rowwise()%>%
  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                                            abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                                            abs(dif_twora), na.rm = T)) %>%
  group_by(unitid) %>%
  summarise(public = mean(public),
            private = mean(private),
            forprofit = mean(forprofit),
            selective = mean(selective),
            more_selective = mean(more_selective),
            non_selective = mean(non_selective),
            avg_magdiff = mean(mag_diff))

four_year_type_avgmagdiff_corr <- four_year_type_magdiff %>%
  select(public, private, forprofit, selective, more_selective, non_selective, avg_magdiff)

```

### Correlation Matrix for Inst. Type and Avg Mag_Diff

```{r code789}
M3 = cor(four_year_type_avgmagdiff_corr)
corrplot(M3, method ='number')
```

### Looking at Correlations for Different Types of Universities


```{r code457}
sixseven<- four_year %>%
  filter(year == 2017) %>%
  rowwise () %>%
  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                        abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                        abs(dif_twora), na.rm = T)) %>%
  group_by(unitid) %>%
  summarise(avgdif_white = mean(dif_white), 
                    avgdif_amind = mean(dif_amind),
                    avgdif_asian = mean(dif_asian),
                    avgdif_black = mean(dif_black),
                    avgdif_pacis = mean(dif_pacis),
                    avgdif_twora = mean(dif_twora),
                    avgdif_hispa = mean(dif_hispa),
                    avgdif_magdif = (mean(mag_diff))) %>%
  select(avgdif_white, 
        avgdif_amind,
        avgdif_asian,
        avgdif_black,
        avgdif_pacis,
        avgdif_twora,
        avgdif_hispa,
        avgdif_magdif)

M4 = cor(sixseven)
corrplot(M4, method ='number', title = "2017 Four-Year Institutions", 
         tl.col="black", mar=c(0,0,1.5,0), tl.offset = 0.5) 
```

### Exercise 1

Insert any text here.

```{r code457}

```