---
title: "Analyzing Ranks after Cuts"
author: "Emily B., Sagma P."
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
#Load needed R Packages
library(tidyverse)
library(openintro)
library(ggplot2)
library(dplyr)
```

### Import, Filter, and Mutate Data

```{r code1}
#Import two year and four year institution dataframes.
two_year <- read_csv("CollegeData_2-year.csv")
four_year <- read_csv("CollegeData_4-year.csv")


#Select only variables needed for metric use in the dataframes. 
four_year_clean <- four_year %>%
                  select(unitid, year, inst_name, dif_white, dif_amind, dif_asian, dif_black,
                        dif_pacis, dif_twora, dif_hispa)
two_year_clean <- two_year %>%
                  select(unitid, year, inst_name, dif_white, dif_amind, dif_asian, dif_black,
                        dif_pacis, dif_twora, dif_hispa)
                        
#Create magnitude difference, and square root of means of squared differences variables in the four year dataframe and two year dataframe.
#Also, create character variable to store number of missing racial data a school has in any given year.
#NOTE: this will take a second to complete, green bar is good!, wait until it disappears 
four_year_md_smsd <- four_year_clean %>%
                  rowwise () %>%
                  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                                            abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                                            abs(dif_twora), na.rm = T)) %>%
                  mutate(sse = sum((dif_white)^2, (dif_hispa)^2, (dif_black)^2, 
                              (dif_asian)^2, (dif_amind)^2, (dif_pacis)^2,
                              (dif_twora)^2, na.rm = T)) %>%
                  mutate(smsd = sqrt(1/7*sse)) %>%
                  mutate(missing_racial_data = is.na(dif_white) + is.na(abs(dif_hispa)) 
                                              + is.na(abs(dif_black)) + is.na(abs(dif_asian)) 
                                              + is.na(abs(dif_amind)) + is.na(abs(dif_pacis))
                                              + is.na(abs(dif_twora))) %>%
                  mutate(missing_racial_data = as.character(missing_racial_data))

two_year_md_smsd <- two_year_clean %>%
                  rowwise () %>%
                  mutate(mag_diff = sum(abs(dif_white), abs(dif_hispa), abs(dif_black),
                                            abs(dif_asian), abs(dif_amind), abs(dif_pacis),
                                            abs(dif_twora), na.rm = T)) %>%
                  mutate(sse = sum((dif_white)^2, (dif_hispa)^2, (dif_black)^2, 
                              (dif_asian)^2, (dif_amind)^2, (dif_pacis)^2,
                              (dif_twora)^2, na.rm = T)) %>%
                  mutate(smsd = sqrt(1/7*sse)) %>%
                  mutate(missing_racial_data = is.na(dif_white) + is.na(abs(dif_hispa)) 
                                              + is.na(abs(dif_black)) + is.na(abs(dif_asian)) 
                                              + is.na(abs(dif_amind)) + is.na(abs(dif_pacis))
                                              + is.na(abs(dif_twora))) %>%
                  mutate(missing_racial_data = as.character(missing_racial_data))
                  
                  
###Cutting out schools with one or more years of data missing
four_year_keeper_list<- four_year_md_smsd %>%
                    group_by(unitid) %>%
                    summarize(count=n()) %>%
                    filter(count == 9) 
two_year_keeper_list<- two_year_md_smsd %>%
                    group_by(unitid) %>%
                    summarize(count=n()) %>%
                    filter(count == 9) 
four_year_md_cut <- four_year_md_smsd %>%
                    filter(unitid %in% four_year_keeper_list$unitid)
two_year_md_cut <- two_year_md_smsd %>%
                    filter(unitid %in% two_year_keeper_list$unitid)
                    
#Determine totoal number of four year and two year schools included in the data befor and after cutting schools misisng one or more years of data. 
length(unique(four_year_md_smsd$unitid))
length(unique(four_year_md_cut$unitid)) 
length(unique(two_year_md_smsd$unitid))
length(unique(two_year_md_cut$unitid)) 

#Crete dataframes to join original dataset to get all needed variables in the metric analysis section.
four_join <- four_year_md_cut %>%
          group_by(unitid) %>%
          filter(row_number()==1)
two_join  <- two_year_md_cut %>%
          group_by(unitid) %>%
          filter(row_number()==1)
```

### Line of Best Fit Using Diff Magnitudes

```{r code2}
#For four year schools, create line of best fit model through magnitude difference values across years for each school.
#Store slope of this line of best fit for each school.
#Join four_join dataframe to include unitid, inst_name, linmod, and slope variables in final data frame.
#Order dataframe rows by slope (ascending). 
four_year_final_md <- four_year_md_cut %>%   
                  group_by(unitid) %>%
                  do(linmod = lm( mag_diff ~ year, data = .)) %>%
                  rowwise () %>%
                  mutate(slope = linmod$coefficients[2]) %>%
                  arrange(slope)
four_year_md_join <- left_join(four_join, four_year_final_md, by=c('unitid'='unitid')) %>%
                  select(unitid, inst_name, linmod, slope)
four_year_md_final <- four_year_md_join %>%
                    arrange(slope)

#For two year schools, create line of best fit model through magnitude difference values across years for each school.
#Store slope of this line of best fit for each school.
#Join four_join dataframe to include unitid, inst_name, linmod, and slope variables in final data frame.
#Order dataframe rows by slope (ascending). 
two_year_final_md <- two_year_md_cut %>%   
                  group_by(unitid) %>%
                  do(linmod = lm( mag_diff ~ year, data = .)) %>%
                  rowwise () %>%
                  mutate(slope = linmod$coefficients[2]) %>%
                  arrange(slope)
two_year_md_join <- left_join(two_join, two_year_final_md, by=c('unitid'='unitid')) %>%
                  select(unitid, inst_name, linmod, slope)
two_year_md_final <- two_year_md_join %>%
                    arrange(slope)
```

### Line of Best Fit Using (sqrt) Mean of Squared Differences 

```{r code3}
#For four year schools, create line of best fit model through (sqrt) Mean of Squared Differences values across years for each school.
#Store slope of this line of best fit for each school.
#Join four_join dataframe to include unitid, inst_name, linmod, and slope variables in final data frame.
#Order dataframe rows by slope (ascending). 
four_year_final_smsd <- four_year_md_cut %>%   
                     group_by(unitid) %>%
                     do(linmod = lm( smsd ~ year, data = .)) %>%
                     rowwise () %>%
                     mutate(slope = linmod$coefficients[2]) %>%
                     arrange(slope)
four_year_smsd_join <- left_join(four_join, four_year_final_smsd, by=c('unitid'='unitid')) %>%
                  select(unitid, inst_name, linmod, slope)
four_year_smsd_final <- four_year_smsd_join %>%
                    arrange(slope)
                    
#For two year schools, create line of best fit model through (sqrt) Mean of Squared Differences values across years for each school.
#Store slope of this line of best fit for each school.
#Join four_join dataframe to include unitid, inst_name, linmod, and slope variables in final data frame.
#Order dataframe rows by slope (ascending). 
two_year_final_smsd <- two_year_md_cut %>%   
                    group_by(unitid) %>%
                    do(linmod = lm( smsd ~ year, data = .)) %>%
                    rowwise () %>%
                    mutate(slope = linmod$coefficients[2]) %>%
                    arrange(slope)
two_year_smsd_join <- left_join(two_join, two_year_final_smsd, by=c('unitid'='unitid')) %>%
                  select(unitid, inst_name, linmod, slope)
two_year_smsd_final <- two_year_smsd_join %>%
                    arrange(slope)
```

### Analyzing common 4 years school

```{r code4}
#For four year schools, create rank_md variable to store rank of md slopes across schools.
#Create rank_smsd variable to store rank of smsd slopes across schools.
four_year_final_md$rank_md<- rank(four_year_final_md$slope)
four_year_final_smsd$rank_smsd <- rank(four_year_final_smsd$slope)

#Join four year final md rank dataframe with four year final smsd dataframe.
common_school <- left_join ( four_year_final_md, four_year_final_smsd, by=c("inst_name"="inst_name"))

#From this larger dataframe select inst_name, rank_md, and rank_smsd variables.
comparing_four_year_school <- common_school %>%
  select(inst_name, rank_md,rank_smsd)

```

### Analyzing common 2 years school
```{r code5}
#For two year schools, create rank_md variable to store rank of md slopes across schools.
#Create rank_smsd variable to store rank of smsd slopes across schools.
two_year_final_md$rank_md<- rank(two_year_final_md$slope)
two_year_final_smsd$rank_smsd <- rank(two_year_final_smsd$slope)

#Join two year final md rank dataframe with two year final smsd dataframe.
common_school_two_year <- left_join ( two_year_final_md, two_year_final_smsd, by=c("inst_name"="inst_name"))

#From this larger dataframe select inst_name, rank_md, and rank_smsd variables.
comparing_two_year_school <- common_school_two_year %>%
  select(inst_name, rank_md,rank_smsd)

```

### Common school plot MD VS SMSD

```{r code6}
#Create dot plot to compare four year schools md ranks and smsd ranks.
plot(comparing_four_year_school$rank_md, comparing_four_year_school$rank_smsd,
     xlab = "MD",
     ylab = "SMSD",
     main = "Rank obtained by MD VS SMSD for 4 year college")
     
#Create dot plot to compare two year schools md ranks and smsd ranks.
plot(comparing_two_year_school$rank_md, comparing_two_year_school$rank_smsd,
     xlab = "MD",
     ylab = "SMSD",
     main = "Rank obtained by MD VS SMSD for 2 year college")

```
### Calculating the difference in the Rank
```{r }
#Calculate the difference in the rankings for each school between md and smsd ranks for both two year and four year schools.
comparing_four_year_school$Difference <- abs (comparing_four_year_school$rank_md - comparing_four_year_school$rank_smsd)
comparing_two_year_school$Difference <- abs (comparing_two_year_school$rank_md - comparing_two_year_school$rank_smsd)
